---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%"
)
```

# txcutr

<!-- badges: start -->
[![R build status](https://github.com/mfansler/txcutr/workflows/R-CMD-check-bioc/badge.svg)](https://github.com/mfansler/txcutr/actions)
[![codecov](https://codecov.io/gh/mfansler/txcutr/branch/main/graph/badge.svg?token=CGGZP68G67)](https://codecov.io/gh/mfansler/txcutr)
[![Anaconda-Server Badge](https://anaconda.org/merv/r-txcutr/badges/installer/conda.svg)](https://conda.anaconda.org/merv/r-txcutr)
[![Anaconda-Server Badge](https://anaconda.org/merv/r-txcutr/badges/version.svg)](https://anaconda.org/merv/r-txcutr)
<!-- badges: end -->

## Overview

Various mRNA sequencing library preparation methods generate sequencing reads 
from the transcript ends. Quantification of isoform usage can be improved by 
using truncated versions of transcriptome annotations when assigning such reads 
to isoforms. The `txcutr` package implements some convenience methods for 
readily generating such truncated annotations and their corresponding sequences.

## Installation instructions

### Bioconductor
Get the latest stable `R` release from [CRAN](http://cran.r-project.org/). Then install `txcutr` using from [Bioconductor](http://bioconductor.org/) the following code:

```{r 'install', eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}

BiocManager::install("txcutr")
```

And the development version from [GitHub](https://github.com/mfansler/txcutr) with:

```{r 'install_dev', eval = FALSE}
BiocManager::install("mfansler/txcutr")
```

### Conda/Mamba

Users managing R environments with Conda/Mamba can install the package with:

**Conda**
```bash
conda install -c conda-forge -c bioconda merv::r-txcutr
```

**Mamba**
```bash
mamba install -c conda-forge -c bioconda merv::r-txcutr
```

We strongly encourage users to create dedicated R environments. **Do not 
install this in your *base* environment!**

## Example

A typical workflow for `txcutr` involves 

 - loading an existing annotation as `TxDb` object
 - truncating the annotation
 - exporting the truncated annotation (GTF)
 - exporting supporting files (FASTA, merge TSV)
 
```{r eval=FALSE}
library(rtracklayer)
library(txcutr)
library(BSgenome.Hsapiens.UCSC.hg38)

## load human genome
hg38 <- BSgenome.Hsapiens.UCSC.hg38

## load human GENCODE annotation
txdb <- makeTxDbFromGFF("gencode.v38.annotaton.gtf.gz", organism="Homo sapiens")

## truncate to maximum of 500 nts
txdb_w500 <- truncateTxome(txdb, maxTxLength=500)

## export annotation
exportGTF(txdb_w500, file="gencode.v38.txcutr_w500.gtf.gz")

## export FASTA
exportFASTA(txdb_w500, genome=hg38, file="gencode.v38.txcutr_w500.fa.gz")

## export merge-table
exportMergeTable(txdb_w500, minDistance=200,
                 file="gencode.v38.txcutr_w500.merge.tsv.gz")
```

## Citation

Below is the citation output from using `citation('txcutr')` in R. Please
run this yourself to check for any updates on how to cite __txcutr__.

```{r 'citation', eval = requireNamespace('txcutr')}
print(citation('txcutr'), bibtex = TRUE)
```

Please note that the `txcutr` was only made possible thanks to many other R and bioinformatics software authors, which are cited either in the vignettes and/or the paper(s) describing this package.

## Code of Conduct

Please note that the `txcutr` project is released with a [Contributor Code of Conduct](http://bioconductor.org/about/code-of-conduct/). By contributing to this project, you agree to abide by its terms.

## Development tools

* Continuous code testing is possible thanks to [GitHub actions](https://www.tidyverse.org/blog/2020/04/usethis-1-6-0/)  through `r BiocStyle::CRANpkg('usethis')`, `r BiocStyle::CRANpkg('remotes')`, and `r BiocStyle::CRANpkg('rcmdcheck')` customized to use [Bioconductor's docker containers](https://www.bioconductor.org/help/docker/) and `r BiocStyle::Biocpkg('BiocCheck')`.
* Code coverage assessment is possible thanks to [codecov](https://codecov.io/gh) and `r BiocStyle::CRANpkg('covr')`.
* The [documentation website](http://mfansler.github.io/txcutr) is automatically updated thanks to `r BiocStyle::CRANpkg('pkgdown')`.
* The code is styled automatically thanks to `r BiocStyle::CRANpkg('styler')`.
* The documentation is formatted thanks to `r BiocStyle::CRANpkg('devtools')` and `r BiocStyle::CRANpkg('roxygen2')`.

For more details, check the `dev` directory.

This package was developed using `r BiocStyle::Biocpkg('biocthis')`.
